# Explore One Variable

## Pseudo-Facebook User Data
  
```{r}
getwd()
list.files()
pf <- read.csv('pseudo_facebook.tsv', sep ='\t')
names(pf)
```

## Histogram of Users' Birthdays

```{r}
library(ggplot2)
names(pf)

qplot(x = dob_day, data = pf) + scale_x_continuous(breaks=1:31)
```

## Faceting

```{r}
qplot(x = dob_day, data = pf) + scale_x_continuous(breaks=1:31) + facet_wrap(~dob_month, ncol = 3)
```
## Friend Count

```{r}
qplot(x = friend_count, data = pf)
```


## Limiting the Axes

```{r}
qplot(x = friend_count, data = pf, xlim = c(0, 1000))
```

## Adjusting the Bin Width

```{r}
qplot(x = friend_count, data = pf, binwidth = 25) + scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50))
```
```{r}
qplot(x = friend_count, data = pf, binwidth = 25) + scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) + facet_wrap(~gender)
```

## Omitting NA Observations
```{r}
qplot(x = friend_count, data = subset(pf, !is.na(gender)), binwidth = 10) + scale_x_continuous(lim = c(0,1000), breaks = seq(0,1000,50)) + facet_wrap(~gender)
```

## Statistics 'by' Gender

```{r}
table(pf$gender)
```
```{r}
by(pf$friend_count, pf$gender, summary)
```

## Tenure

```{r}
qplot(x = tenure, data = pf, color = I('black'), fill = I('#099DD9'))
```
```{r}
qplot(x = tenure, data = pf, binwidth=30, color = I('black'), fill = I('#099DD9'))
```
```{r}
qplot(x = tenure/365, data = pf, binwidth=.25, color = I('black'), fill = I('#F79420'))
```
```{r}
qplot(x = tenure/365, data = pf, binwidth=.25, color = I('black'), fill = I('#F79420')) + scale_x_continuous(breaks = seq(1,7,1), limits = c(0,7))
```

## Labelling Plots
```{r}
qplot(x = tenure/365, data = pf, xlab = 'Number of years using Facebook', ylab = 'Number of users in sample', binwidth=.25, color = I('black'), fill = I('#F79420')) + scale_x_continuous(breaks = seq(1,7,1), limits = c(0,7))
```
## User Ages

```{r}
qplot(x = age, data = pf, color = I('black'), fill = I('#5760AB'))
```
```{r}
qplot(x = age, data = pf, binwidth = 1, color = I('black'), fill = I('#5760AB'))
```
```{r}
qplot(x = age, data = pf, binwidth = 1, color = I('black'), fill = I('#5760AB')) + scale_x_continuous(breaks = seq(0, 113, 5))
```
## Transforming Data

```{r}
qplot(x = friend_count, data = pf)

summary(pf$friend_count)

summary(log10(pf$friend_count + 1))

summary(sqrt(pf$friend_count))
```
```{r}
install.packages('gridExtra')
library(gridExtra)
```

```{r}
p1 <- qplot(x = friend_count, data = pf)
p2 <- qplot(x = log10(friend_count + 1), data = pf)
p3 <- qplot(x = sqrt(friend_count), data = pf)
```

```{r}
grid.arrange(p1, p2, p3, ncol = 1)
```
```{r}
p1 <- ggplot(aes(x = friend_count), data = pf) + geom_histogram()
p2 <- p1 + scale_x_log10()
p3 <- p1 + scale_x_sqrt()

grid.arrange(p1, p2, p3, ncol = 1)
```
## Add a Scaling Layer

```{r}
logScale <- qplot(x = log10(friend_count), data = pf)

countScale <- ggplot(aes(x = friend_count), data = pf) + geom_histogram() + scale_x_log10()

grid.arrange(logScale, countScale, ncol = 2)
```
## Frequency Polygons

```{r}
qplot(x = friend_count, data = subset(pf, !is.na(gender)), binwidth = 10) + scale_x_continuous(lim = c(0,1000), breaks = seq(0, 1000, 50)) + facet_wrap(~gender)
```
```{r}
qplot(x = friend_count, data = subset(pf, !is.na(gender)), binwidth = 10, geom = 'freqpoly', color = gender) + scale_x_continuous(lim = c(0,1000), breaks = seq(0, 1000, 50))
```
```{r}
qplot(x = friend_count, y = ..count../sum(..count..) ,data = subset(pf, !is.na(gender)), xlab = 'Friend Count', ylab = 'Propportion of Users with that friend count', binwidth = 10, geom = 'freqpoly', color = gender) + scale_x_continuous(lim = c(0,1000), breaks = seq(0, 1000, 50))
```

```{r}
qplot(x = www_likes, data = subset(pf, !is.na(gender)), geom = 'freqpoly')

qplot(x = www_likes, data = subset(pf, !is.na(gender)), geom = 'freqpoly', color = gender)

qplot(x = www_likes, data = subset(pf, !is.na(gender)), geom = 'freqpoly', color = gender) + scale_x_continuous() + scale_x_log10()
```
## Likes on the Web

```{r}
by(pf$www_likes, pf$gender, sum)
```
## Box Plots

```{r}
qplot(x = gender, y = friend_count, data = subset(pf, !is.na(gender)), geom = 'boxplot')
```
```{r}
qplot(x = gender, y = friend_count, data = subset(pf, !is.na(gender)), geom = 'boxplot') + scale_y_continuous(limits = c(0, 1000))
```

```{r}
qplot(x = gender, y = friend_count, data = subset(pf, !is.na(gender)), geom = 'boxplot') + coord_cartesian(ylim = c(0,1000))
```
## Box Plots, Quartiles and Friendships
```{r}
qplot(x = gender, y = friend_count, data = subset(pf, !is.na(gender)), geom = 'boxplot') + coord_cartesian(ylim = c(0,250))

by(pf$friend_count, pf$gender, summary)
```
```{r}
qplot(x = gender, y = friendships_initiated, data = subset(pf, !is.na(gender)), geom = 'boxplot') + coord_cartesian(ylim = c(0,500))
```
```{r}
qplot(x = gender, y = friendships_initiated, data = subset(pf, !is.na(gender)), geom = 'boxplot') + coord_cartesian(ylim = c(0,150))

by(pf$friendships_initiated, pf$gender, summary)
```
## Getting Logical
```{r}
summary(pf$mobile_likes > 0)

mobile_check_in <- NA
pf$mobile_check_in <- ifelse(pf$mobile_likes > 0, 1, 0)
pf$mobile_check_in <- ifelse(pf$mobile_likes)
summary(pf$mobile_check_in)
```

```{r}
summary(pf$mobile_check_in)
sum(pf$mobile_check_in == 1)/length(pf$mobile_check_in)
```