# Diamonds & Price Predictions

## Scatterplot review

```{r}
diamonds <- read.csv('diamondsbig.csv', sep = '\t')
library('ggplot2')
data(diamonds)
```

```{r}
qplot(data = diamonds, x = carat, y = price, xlim = c(0, quantile(diamonds$carat, 0.99)), ylim = c(0, quantile(diamonds$price, 0.99))) + geom_point(fill = I('#F79420'), color = I('black'), shape = 21)
```

## Price and Carat Relationship
```{r}
ggplot(data = diamonds, aes(x = carat, y = price)) + geom_point(color = '#F79420', alpha = 1/4) + stat_smooth(method = 'lm') + scale_x_continuous(lim = c(0, quantile(diamonds$carat, 0.99))) + scale_y_continuous(lim = c(0, quantile(diamonds$price, 0.99)))
```
## ggpairs()
```{r}
# install these if necessary
install.packages('GGally')
install.packages('scales')
install.packages('memisc')
install.packages('lattice')
install.packages('MASS')
install.packages('car')
install.packages('reshape')
install.packages('plyr')
```

```{r}
# load the ggplot graphics package and the others
library(ggplot2)
library(GGally)
library(scales)
library(memisc)
```
```{r}
# sample 10,000 diamonds from the data set
set.seed(20022012)
diamond_samp <- diamonds[sample(1:length(diamonds$price), 10000), ]
ggpairs(diamond_samp, shape = c(shape = I('.'), outlier.shape = I('.')))
```
## The Demand of Diamonds 
```{r}
plot1 <- qplot(data = diamonds, x = price, binwidth = 100, fill = I('#099DD9')) + ggtitle('Price')

plot2 <- qplot(data = diamonds, x = price, binwidth = 0.01, fill = I('#F79420')) + ggtitle('Price (log10)') + scale_x_log10()

library(gridExtra)
library(grid)
grid.arrange(plot1, plot2, ncol = 2)
```
## Scatterplot Transformation
```{r}
qplot(carat, price, data = diamonds) + scale_y_continuous(trans = log10_trans()) + ggtitle('Price (log10) by Carat')
```

### Create a new function to transform the carat variable
```{r}
cuberoot_trans = function() trans_new('cuberoot', transform = function(x) x^(1/3), inverse = function(x) x^3)
```

### Use the cuberoot_trans function
```{r}
ggplot(aes(carat, price), data = diamonds) + geom_point() + scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3), breaks = c(0.2, 0.5, 1, 2, 3)) + scale_y_continuous(trans = log10_trans(), limits = c(350, 15000), breaks = c(350, 1000, 5000, 10000, 15000)) + ggtitle('Price (log10) by Cube-Root of Carat')
```
## Overplotting Revisited
```{r}
ggplot(aes(carat, price), data = diamonds) + geom_point(alpha = 0.5, size = 0.75, position = 'jitter') + scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3), breaks = c(0.2, 0.5, 1, 2, 3)) + scale_y_continuous(trans = log10_trans(), limits = c(350, 15000), breaks = c(350, 1000, 5000, 10000, 15000)) + ggtitle('Price (log10) by Cube-Root of Carat')
```

## Price vs. Carat and Clarity 
```{r}
install.packages('RColorBrewer', dependencies = TRUE)
library(RColorBrewer)

ggplot(aes(x = carat, y = price, colour = clarity), data = diamonds) + geom_point(alpha = 0.5, size = 1, position = 'jitter') + scale_color_brewer(type = 'div', guide = guide_legend(title = 'Clarity', reverse = TRUE, override.aes = list(alpha = 1, size = 2))) + scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3), breaks = c(0.2, 0.5, 1, 2, 3)) + scale_y_continuous(trans = log10_trans(), limits = c(350, 15000), breaks = c(350, 1000, 5000, 10000, 15000)) + ggtitle('Price (log10) vs Cube Root of Carat and Clarity')
```
## Price vs Carat and Cut
```{r}
ggplot(aes(x = carat, y = price, color = cut), data = diamonds) + geom_point(alpha = 0.5, size = 1, position = 'jitter') + scale_color_brewer(type = 'div', guide = guide_legend(title = 'Cut', reverse = TRUE, override.aes = list(alpha = 1, size = 2))) + scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3), breaks = c(0.2, 0.5, 1, 2, 3)) + scale_y_continuous(trans = log10_trans(), limits = c(350, 15000), breaks = c(350, 1000, 5000, 10000, 15000)) + ggtitle('Price (log10) vs Cube Root of Carat and Cut')
```

## Price vs Carat and Color
```{r}
ggplot(aes(x = carat, y = price, color = color), data = diamonds) + geom_point(alpha = 0.5, size = 1, position = 'jitter') + scale_color_brewer(type = 'div', guide = guide_legend(title = 'Color', reverse = FALSE, override.aes = list(alpha = 1, size = 2))) + scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3), breaks = c(0.2, 0.5, 1, 2, 3)) + scale_y_continuous(trans = log10_trans(), limits = c(350, 15000), breaks = c(350, 1000, 5000, 10000, 15000)) + ggtitle('Price (log10) vs Cube Root of Carat and Color')
```
## Building the linear model
```{r}
m1 <- lm(I(log(price)) ~ I(carat^(1/3)), data = diamonds)
m2 <- update(m1, ~ . + carat)
m3 <- update(m2, ~ . + cut)
m4 <- update(m3, ~ . + color)
m5 <- update(m4, ~ . + clarity)
mtable(m1, m2, m3, m4, m5, sdigits = 3)
```
##
```{r}
#install.packages('RCurl')
#install.packages('bitops')
library('RCurl')
library('bitops')
diamondsurl = getBinaryURL("https://raw.github.com/solomonm/diamonds-data/master/BigDiamonds.Rda")
load(rawConnection(diamondsurl))
```

```{r}
library(readr)
diamondsbig <- read_csv("diamondsbig.csv")
```

```{r}
diamondsbig$logprice = log(diamondsbig$price)
m1 <- lm(logprice ~ I(carat^(1/3)), data=diamondsbig[diamondsbig$price < 10000 & diamondsbig$cert == "GIA",])
m2 <- update(m1, ~ . + carat)
m3 <- update(m2, ~ . + cut)
m4 <- update(m3, ~ . + color)
m5 <- update(m4, ~ . + clarity)
#library(memisc)
mtable(m1, m2, m3, m4, m5)
```

## Predictions

```{r}
thisDiamond = data.frame(carat = 1.00, cut = "V.Good", color = "I", clarity="VS1")
modelEstimate = predict(m5, newdata = thisDiamond, interval="prediction", level = .95)

exp(modelEstimate)

dat = data.frame(m4$model, m4$residuals)

with(dat, sd(m4.residuals))

with(subset(dat, carat > .9 & carat < 1.1), sd(m4.residuals))

dat$resid <- as.numeric(dat$m4.residuals)
ggplot(aes(y = resid, x = round(carat, 2)), data = dat) +
  geom_line(stat = "summary", fun.y = sd)
```