# Lesson 5

## Scatter Plots
```{r}
library(ggplot2)

pf <- read.csv('pseudo_facebook.tsv', sep = '\t')

qplot(x = age, y = friend_count, data = pf)
qplot(age, friend_coint, data = pf)
```

## ggplot Syntax
```{r}
ggplot(aes(x = age, y = friend_count), data = pf) + geom_point()
summary(pf$age)

ggplot(aes(x = age, y = friend_count), data = pf) + geom_point() + xlim(13,90)
```

## Overplotting
```{r}
ggplot(aes(x = age, y = friend_count), data = pf) + geom_jitter(alpha = 1/20) + xlim(13,90)
```

## Coord_trans()
```{r}
ggplot(aes(x = age, y = friend_count), data = pf) + geom_point(alpha = 1/20, position = position_jitter(h = 0)) + xlim(13,90) + coord_trans(y = 'sqrt')
```
## Alpha and Jitter
```{r}
ggplot(aes(x = age, y = friendships_initiated), data = pf) + geom_point(alpha = 1/10, position = position_jitter(h = 0)) + coord_trans(y = 'sqrt')
```
## Conditional Means
```{r}
install.packages('dplyr')
library(dplyr)

age_groups <- group_by(pf, age)
pf.fc_by_age <- summarise(age_groups, friend_count_mean = mean(friend_count), friend_count_median = median(friend_count), n = n())
pf.fc_by_age <- arrange(pf.fc_by_age, age)
head(pf.fc_by_age)
```
```{r}
pf.fc_by_age <- pf %>%
	group_by(age) %>%
	summarise(friend_count_mean = mean(friend_count), friend_count_median = median(friend_count), n = n()) %>%
	arrange(age)

head(pf.fc_by_age, 20)
```
```{r}
library(ggplot2)

ggplot(aes(age, friend_count_mean), data = pf.fc_by_age) + geom_point()

ggplot(aes(age, friend_count_mean), data = pf.fc_by_age) + geom_line()
```
## Overlaying Summaries with Raw Data
```{r}
ggplot(aes(x = age, y = friend_count), data = pf) + xlim(13, 90) + geom_point(alpha = 0.05, position = position_jitter(h=0), color = 'orange')+ coord_trans(y = 'sqrt')

ggplot(aes(x = age, y = friend_count), data = pf) + xlim(13, 90) + geom_point(alpha = 0.05, position = position_jitter(h=0), color = 'orange')+ coord_trans(y = 'sqrt') + geom_line(stat = 'summary', fun.y = mean)

ggplot(aes(x = age, y = friend_count), data = pf) + xlim(13, 90) + geom_point(alpha = 0.05, position = position_jitter(h=0), color = 'orange')+ coord_trans(y = 'sqrt') + geom_line(stat = 'summary', fun.y = mean) + geom_line(stat = 'summary', fun.y = quantile, fun.args = list(probs = .1), linetype = 2 , color = 'blue')

ggplot(aes(x = age, y = friend_count), data = pf) + xlim(13, 90) + geom_point(alpha = 0.05, position = position_jitter(h=0), color = 'orange')+ coord_trans(y = 'sqrt') + geom_line(stat = 'summary', fun.y = mean) + geom_line(stat = 'summary', fun.y = quantile, fun.args = list(probs = .1), linetype = 2 , color = 'blue') + geom_line(stat = 'summary', fun.y = quantile, fun.args = list(probs = .9), linetype = 2 , color = 'blue')
```

## Correlation
```{r}
cor.test(pf$age, pf$friend_count, method = 'pearson')

with(pf, cor.test(age, friend_count, method = 'pearson'))
```
## Correlation on Subsets
```{r}
with(subset(pf, age < 70), cor.test(age, friend_count))
with(subset(pf, age < 70), cor.test(age, friend_count, method = 'pearson'))
```
## Correlation Methods
```{r}
with(subset(pf, age < 70), cor.test(age, friend_count, method = 'spearman'))
```

## Create Scatterplots
```{r}
ggplot(aes(x = www_likes_received, y = likes_received), data = pf) + geom_point()
```

## Strong Correlation
```{r}
gplot(aes(x = www_likes_received, y = likes_received), data = pf) + geom_point() + xlim(0, quantile(pf$www_likes_received, 0.95)) + ylim(0, quantile(pf$likes_received, 0.95))

ggplot(aes(x = www_likes_received, y = likes_received), data = pf) + geom_point() + xlim(0, quantile(pf$www_likes_received, 0.95)) + ylim(0, quantile(pf$likes_received, 0.95)) + geom_smooth(method = 'lm', color ='red')

cor.test(pf$www_likes_received, pf$likes_received)
```
## More Caution with Correlation
```{r}
install.packages('alr3')
library(alr3)
data(Mitchell)
?Mitchell
```
### Create a Scatterplot of Temperature vs. Months
```{r}
ggplot(data = Mitchell, aes(x = Month, y = Temp)) + geom_point()
qplot(data = Mitchell, Month, Temp)
```
## Noisy Scatterplots
```{r}
cor.test(Mitchell$Month, Mitchell$Temp)
```
## Making Sense of Data
```{r}
ggplot(data = Mitchell, aes(x = Month, y = Temp)) + geom_point() + scale_x_discrete(breaks = seq(0, 203, 12))
```
## Understanding Noise: Age to Age Months
```{r}
ggplot(aes(x = age, y = friend_count_mean), data = pf.fc_by_age) + geom_line()

head(pf.fc_by_age, 10)
pf.fc_by_age[17:19, ]
```
```{r}
pf$age_with_months <- pf$age + (12 - pf$dob_month) / 12
```
## Age with month means
```{r}
library(dplyr)
pf.fc_by_age_months <- pf %>%
	group_by(age_with_months) %>%
	summarise(friend_count_mean = mean(friend_count), friend_count_median = median(friend_count), n = n()) %>%
arrange(age_with_months)

head(pf.fc_by_age_months)
```
```{r}
age_with_months_groups <- group_by(pf, age_with_months)
pf.fc_by_age_months2 <- summarise(age_with_months_groups, friend_count_mean = mean(friend_count), friend_count_median = median(friend_count), n = n())
pf.fc_by_age_months2 <- arrange(pf.fc_by_age_months2, age_with_months)

head(pf.fc_by_age_months2)
```
## Noise in Conditional Means
```{r}
ggplot(aes(x = age_with_months, y = friend_count_mean), data = subset(pf.fc_by_age_months, age_with_months < 71)) + geom_line()
```
## Smoothing Conditinoal Means
```{r}
p1 <- ggplot(aes(x = age, y = friend_count_mean), data = pf.fc_by_age, age < 71) + geom_line() + geom_smooth()

p2 <- ggplot(aes(x = age_with_months, y = friend_count_mean), data = subset(pf.fc_by_age_months, age_with_months < 71)) + geom_line() + geom_smooth()

p3 <- ggplot(aes(x = round(age / 5) * 5, y = friend_count), data = subset(pf, age < 71)) + geom_line(stat = 'summary', fun.y = mean)

library(gridExtra)
grid.arrange(p2, p1, p3, ncol = 1)
```

##

##