# Explore Many Variables

```{r}
library(ggplot2)

pf <- read.csv('pseudo_facebook.tsv', sep = '\t')
```

## Third Qualitative Variable

```{r}
ggplot(aes(x = gender, y = age), data = subset(pf, !is.na(gender))) + geom_boxplot() + stat_summary(fun.y = mean, geom = 'point', shape = 4)

ggplot(aes(x = age, y = friend_count), data = subset(pf, !is.na(gender))) + geom_line(aes(color = gender), stat = 'summary', fun.y = median)
```

```{r}
library(dplyr)
# chain functions %>%
pf.fc_by_age_gender <- pf %>%
	group_by(age,gender) %>%
	summarise(mean_friend_count = mean(friend_count), median_friend_count = median(friend_count), n = n()) %>%
	ungroup() %>%
	arrange(age)
```

## Plotting Conditional Summaries
```{r}
ggplot(aes(x = age, y = median_friend_count), data = pf.fc_by_age_gender) + geom_line(aes(color = gender))
```
### Wide and Long Format
```{r}
pf.fc_by_age <- pf %>%
	group_by(age) %>%
	summarise(friend_count_mean = mean(friend_count), friend_count_median = median(friend_count), n = n()) %>%
	arrange(age)

head(pf.fc_by_age, 11)
```
## Reshaping in Data
```{r}
install.packages('reshape2')
library(reshape2)
```
```{r}
pf.fc_by_age_gender.wide <- dcast(pf.fc_by_age_gender, age ~ gender, value.var = 'median_friend_count')

head(pf.fc_by_age_gender.wide)
```
## Ratio Plot

```{r}
ggplot(aes(x = age, y = female / male), data = pf.fc_by_age_gender.wide) + geom_line() + geom_hline(yintercept = 1, alpha = 0.3, linetype = 2)
```

## Third Quantitative Variable
```{r}
pf$year_joined <- floor(2014 - pf$tenure/365)
```

## Cut a Variable
```{r}
summary(pf$year_joined)
table(pf$year_joined)
```
```{r}
pf$year_joined.bucket <- cut(pf$year_joined, c(2004, 2009, 2011, 2012, 2014))
```

## Plotting It All Together
```{r}
table(pf$year_joined.bucket, useNA = 'ifany')

ggplot(aes(x = age, y = friend_count), data = subset(pf, !is.na(gender))) + geom_line(aes(color = gender), stat = 'summary', fun.y = median)

ggplot(aes(x = age, y = friend_count), data = subset(pf, !is.na(year_joined.bucket))) + geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = median)
```
## Plot the Grand Means
```{r}
ggplot(aes(x = age, y = friend_count), data = subset(pf, !is.na(year_joined.bucket))) + geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean) + geom_line(stat = 'summary', fun.y = mean, linetype = 2)
```
## Friending Rate
```{r}
with(subset(pf, tenure >= 1), summary(friend_count / tenure))
```
## Friendships Initiated
```{r}
ggplot(aes(x = tenure, y = friendships_initiated / tenure), data = subset(pf, tenure >= 1)) + geom_line(aes(color = year_joined.bucket))
```
## Bias Variance Trade off Revisited
```{r}
ggplot(aes(x = tenure, y = friendships_initiated / tenure), data = subset(pf, tenure >= 1)) + geom_smooth(aes(color = year_joined.bucket))
```
# Yogurt
```{r}
yo <- read.csv('yogurt.csv')
str(yo)
```

```{r}
# Change the id from an int to a factor
yo$id <- factor(yo$id)
str(yo)
```
```{r}
qplot(data = yo, x = price, fill = I('#F79420'))

qplot(data = yo, x = price, fill = I('#F79420'), binwidth = 10)
```
## Number of Purchases
```{r}
summary(yo)
unique(yo$price)
length(unique(yo$price))
table(yo$price)
```
```{r}
yo <- transform(yo, all.purchases = strawberry + blueberry + pina.colada + plain + mixed.berry)
```

## Prices over Time
```{r}
qplot(x = all.purchases, data = yo, binwidth = 1, fill = I('#099DD9'))
```
```{r}
ggplot(aes(x = time, y = price), data = yo) + geom_jitter(alpha = 1/4, shape = 21, fill = I('#F79420'))
```
## Looking at Samples of Households
```{r}
set.seed(4230)
sample.ids <- sample(levels(yo$id), 16)
```

```{r}
ggplot(aes(x = time, y = price), data = subset(yo, id %in% sample.ids)) + facet_wrap( ~ id) + geom_line() + geom_point(aes(size = all.purchases), pch = 1)
```
## Scatterplots Matrices
```{r}
#install.packages(`GGally`)
library(GGally)
theme_set(theme_minimal(20))
```

```{r}
# set the seed for reproducible results
set.seed(1836)
pf_subset <- pf[, c(2:15)]
names(pf_subset)
ggpairs(pf_subset[sample.int(nrow(pf_subset), 1000), ])
```

## Even More Variables
```{r}
nci <- read.table('nci.tsv')

#changing the colnames to produce a nicer plot
colnames(nci) <- c(1:64)
```

## Heat Map
```{r}
# melt the data to long format
library(reshape2)
nci.long.samp <- melt(as.matrix(nci[1:200, ]))
names(nci.long.samp) <- c('gene', 'case', 'value')
head(nci.long.samp)

# make the heat map
ggplot(aes(y = gene, x = case, fill = value), data = nci.long.samp) + geom_tile() + scale_fill_gradientn(colors = colorRampPalette(c('blue', 'red'))(100))
```